Experiences with Planning for Natural Language Generation

### 2009-06-03 trunk/Papers/ci-spark/experiments.tex
changes to section 4

\subsection{Experiment 2: Minimal GIVE worlds}
\label{sec:exper-2:-minim}

In the second set of experiments, we evaluate the performance of several
planners on problems arising in the GIVE domain. Besides the FF, Metric-FF, and
SGPLAN planners we used in the previous set of experiments, we expand our
testing to include FF(h$_a$), LAMA, and C$^3$ \todo{citations}. Each of these
additional planners competed in the deterministic ``sequential, satifying''
track of the 2008 International Planning Competition; all planners generally
performed well on the competition domains, with LAMA the overall winner of
the track.\footnote{See \texttt{http://ipc.informatik.uni-freiburg.de/} for
details of the 2008 IPC.}

In the first GIVE test, we construct a series of grid worlds, similar to the one
illustrated in Figure~\ref{fig:give-maps}(a). These worlds consist of a $2n$ by
$h$ grid of positions, such that there are buttons at positions $(2i-1,1)$ and
$(2i,h)$ for $1 \leq i \leq n$. The player starts in position $(1,1)$ and must
press all the buttons to successfully complete the game.  (The actions in this
domain are similar to the PDDL actions in Figure~\ref{fig:give-planning}.)
We consider two variants of this problem in our tests. In the \emph{unordered}
problem, the player is permitted to press the buttons in any order to
successfully achieve the goal. In the \emph{ordered} version of the problem, the
player is unable to initially move to any grid cell containing a button, except
for the cell containing the first button, $u_1$. Pressing $u_1$ releases the
position of the next button, $l_1$, allowing the player to move into this cell.
Similarly, pressing button $l_1$ frees button $u_2$, and so on. The end result
is a set of constraints that forces the buttons to be pressed in a particular
order to achieve the goal. As a concrete example, the following is a minimal
plan (in either variant of the problem) for the case of a $2$ by $2$ grid with
$2$ buttons (i.e., $n=1$, $h=2$):

\begin{enumerate}
\item $\mathsf{move}(\mathsf{pos\_1\_1},\mathsf{pos\_1\_2}, \mathsf{north})$,
\item $\mathsf{manipulate}\textsf{-}\mathsf{u1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_1\_2})$,
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{north}, \mathsf{east})$,
\item $\mathsf{move}(\mathsf{pos\_1\_2}, \mathsf{pos\_2\_2},
  \mathsf{east})$,
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{east}, \mathsf{south})$,
\item $\mathsf{move}(\mathsf{pos\_2\_2}, \mathsf{pos\_2\_1}, \mathsf{south})$,
\item $\mathsf{manipulate}\textsf{-}\mathsf{l1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_2\_1})$.
\end{enumerate}

Results for the $h=20$ case, with the grid width $N$ ranging from $0$ to $40$,
are shown in Figure~\ref{fig:give-minimal}. In the unordered case
(Figure~\ref{fig-give-minimal}(a)), the most obvious result is that some of the
planners tested---Metric-FF, FF(h$_a$), and C$^3$---are unable to solve any
problems beyond $N=24$ on our experimentation machine within the memory limit of
2 GB.\footnote{\todo{describe testing environment}} While FF, Lama, and SGPLAN
are all able to solve problem instances up to $N=40$, the total running time
varies greatly between this planners. For instance, FF takes almost 35
seconds to solve the $N=40$ problem, LAMA takes around 6.5 seconds, and SGPLAN
shows impressive performance, taking well less than a second.

In a real NLG system, where response time is

machine within the memory limit of 1 GB. SGPLAN, on the other hand, solves
instances beyond $n=40$ without major problems. In this case, our
implementation of GraphPlan is unable to find any of these plans. The time
spent on grounding is not a major factor in either planner, probably
because the planners need more time to actually compute the plan---for
instance, the optimal plan for the problem $n=40$ has a length of about
1600 steps.

\begin{figure}[t]
\subfloat[(a) Unordered]{\includegraphics[width=7cm]{data/give-minimal.pdf}}
\subfloat[(b) Ordered]{\includegraphics[width=7cm]{data/give2-minimal.pdf}}
  \caption{Results for the unordered and ordered minimal GIVE domains with grid
  height $h=20$. The horizontal axis is the grid width, $N = 2n$. The
  vertical axis is the total runtime in seconds.}
  \label{fig:give-minimal}
\end{figure}

\begin{figure}[t]
\subfloat[(a) Unordered]{\includegraphics[width=7cm]{data/give-minimal-grounding.pdf}}
\subfloat[(b) Ordered]{\includegraphics[width=7cm]{data/give2-minimal-grounding.pdf}}
  \caption{Comparison of the total running time and grounding time for selected
  planners in the $h=20$ minimal GIVE domain. The horizontal axis is the grid
  width, $N = 2n$. The vertical axis is the total runtime in seconds.}
  \label{fig:give-minimal-grounding}
\end{figure}

\subsection{Experiment 3: GIVE worlds with extra positions}
\label{sec:experiment-3:-give}

@@

\subsection{Experiment 4: GIVE worlds with inaccessible positions}
\label{sec:experiment-4:-give}

>>>>

\subsection{Experiment 3: Minimal GIVE worlds}
\label{sec:exper-2:-minim}

We now turn our attention to a set of experiments arising from the GIVE domain.
Besides using some of the planners from the previous set of experiments (FF,
Metric-FF, and SGPLAN), we expand our testing to include the FF(h$_a$), LAMA, and
C$^3$ planners \todo{citations}. Each of these additional planners competed in the
deterministic ``sequential, satifying'' track of the 2008 International Planning
Competition; all planners generally performed well on the competition domains,
with LAMA the overall winner of the track.\footnote{See
\texttt{http://ipc.informatik.uni-freiburg.de/} for details of the 2008 IPC.}

In the first GIVE experiment, we construct a series of grid worlds, similar to
the one illustrated in Figure~\ref{fig:give-maps}(a). These worlds consist of a
$N = 2n$ by $h$ grid of positions, such that there are buttons at positions
$(2i-1,1)$ and $(2i,h)$ for $1 \leq i \leq n$. The player starts in position
$(1,1)$ and must press all the buttons to successfully complete the game.  (The
actions in this domain are similar to the PDDL actions in
Figure~\ref{fig:give-planning}.) We consider two variants of this problem in our
tests. In the \emph{unordered} problem, the player is permitted to press the
buttons in any order to successfully achieve the goal. In the \emph{ordered}
version of the problem, the player is unable to initially move to any grid cell
containing a button, except for the cell containing the first button, $u_1$.
Pressing $u_1$ releases the position of the next button, $l_1$, allowing the
player to move into this cell.  Similarly, pressing button $l_1$ frees button
$u_2$, and so on. The end result is a set of constraints that forces the buttons
to be pressed in a particular order to achieve the goal. As a concrete example,
the following is a minimal plan (in either variant of the problem) for the case
of a $2$ by $2$ grid with $2$ buttons (i.e., $n=1$, $h=2$):

\begin{enumerate}
\item $\mathsf{move}(\mathsf{pos\_1\_1},\mathsf{pos\_1\_2}, \mathsf{north})$,
\item $\mathsf{manipulate}\textsf{-}\mathsf{button}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{u1, pos\_1\_2})$,
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{north}, \mathsf{east})$,
\item $\mathsf{move}(\mathsf{pos\_1\_2}, \mathsf{pos\_2\_2},
  \mathsf{east})$,
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{east}, \mathsf{south})$,
\item $\mathsf{move}(\mathsf{pos\_2\_2}, \mathsf{pos\_2\_1}, \mathsf{south})$,
\item $\mathsf{manipulate}\textsf{-}\mathsf{button}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{l1, pos\_2\_1})$.
\end{enumerate}

Results for the $h=20$ case, with the grid width $N$ ranging from $1$ to $40$,
are shown in Figure~\ref{fig:give-minimal}. In the unordered case
(Figure~\ref{fig:give-minimal}(a)), the most obvious result is that some of the
planners tested---Metric-FF, FF(h$_a$), and C$^3$---are unable to solve any
problems beyond $N=24$ on our experimentation machine within the memory limit of
2 GB.\footnote{\todo{describe testing environment}} While FF, LAMA, and SGPLAN
are able to solve all problem instances up to $N=40$, the total running time
varies greatly between these planners. For instance, FF takes almost 35 seconds
to solve the $N=40$ problem, while LAMA takes around 6.5 seconds. SGPLAN shows
impressive performance, generating a 240 step plan in the $N=40$ case in well
less than a second. In the ordered case (Figure~\ref{fig:give-minimal}(b)), we
again have the situation where Metric-FF, FF(h$_a$), and C$^3$ are unable to
solve all problem instances. Furthermore, both SGPLAN and LAMA, which performed
well on the unordered problem, now perform much worse than FF: FF takes 39
seconds for the $N=40$ case, while SGPLAN takes 50 seconds and LAMA takes 90
seconds. In real NLG systems, where response time is essential, running times
over a few seconds are unacceptable.

Preprocessing (parsing, grounding, etc.) time generally plays less of a role in
GIVE, compared with the sentence generation domain, however, its effects still
contribute significantly to the overall running time of a number of planners.
Figure~\ref{fig:give-minimal-grounding} shows the grounding time for FF, LAMA,
and SGPLAN on the minimal GIVE problems, compared with the total running time.
In the unordered variant of the minimal GIVE domain
(Figure~\ref{fig:give-minimal-grounding}(a)), the grounding time in LAMA and
SGPLAN accounts for a significant fraction of the total runtime: SGPLAN spends
around 40\% of its total runtime on preprocessing; for LAMA, this number
rises to at least 80\% for our test problems. For FF, the preprocessing time is
much less important than the search time, especially for large problem
instances. In the ordered case (Figure~\ref{fig:give-minimal-grounding}(b)), the
actual time spent on preprocessing is essentially unchanged from the unordered
case, and search time dominates the total runtime for all three planners.
However, FF is now much better at controlling the search,
compared with the unordered variant of the problem.

\begin{figure}[t]
\subfloat[(a) Unordered]{\includegraphics[width=7cm]{data/give-minimal.pdf}}
\subfloat[(b) Ordered]{\includegraphics[width=7cm]{data/give2-minimal.pdf}}
  \caption{Results for the unordered and ordered minimal GIVE domains with grid
  height $h=20$. The horizontal axis is the grid width, $N$. The
  vertical axis is the total runtime in seconds.}
  \label{fig:give-minimal}
\end{figure}

\begin{figure}[t]
\subfloat[(a) Unordered]{\includegraphics[width=7cm]{data/give-minimal-grounding.pdf}}
\subfloat[(b) Ordered]{\includegraphics[width=7cm]{data/give2-minimal-grounding.pdf}}
  \caption{Comparison of the total running time and grounding time for selected
  planners in the $h=20$ minimal GIVE domain. The horizontal axis is the grid
  width, $N$. The vertical axis is the total runtime in seconds.}
  \label{fig:give-minimal-grounding}
\end{figure}

\subsection{Experiment 4: GIVE worlds with extra positions}
\label{sec:experiment-3:-give}

@@

\subsection{Experiment 5: GIVE worlds with inaccessible positions}
\label{sec:experiment-4:-give}
